Geometrization of characters of tori over non-Archimedean local fields

### 2013-08-14 shanghai.tex
Working on the last section on transfer ....

\begin{theorem}\label{thm:application}
Let $d$ be a positive integer.
\begin{enumerate}
\item 
For every algebraic torus $T$ over $K$, the trace of Frobenius 
  \[
\trFrob{T,d}:  \QCiso{\GN{T}_d} \to \Hom_{< d}(T(K),\EEx)
  \]
is an isomomorphism. 
\item The isomorphism $\trFrob{T,d}$ is functorial in the following sense: $T \mapsto \trFrob{T,d}$ is a natural transformation
from the additive functor
\[T \mapsto \QCiso{\GN{T}_d}\]
to the additive functor
\[
T \mapsto \Hom_{<d}(T(K),\EEx).
\]
\item
For every finite Galois extension $L/K$ and for every algebraic torus $T$ over $K$, the following diagram commutes:
\[
\begin{tikzcd}
\QCiso{\Gr(\underline{\Res_{L/K}(T_L)}_d)} \arrow[two heads]{r} \dar & \QCiso{\Gr({T}_d)} \dar\\
\Hom_{<d}(T(L), \EEx) \arrow[two heads]{r} & \Hom_{<d}(T(K), \EEx)
\end{tikzcd}
\]
\end{enumerate}
\end{theorem}

@@

\begin{remark}
The Greenberg transform of the N\'eron model of $T$ is a commutative group scheme over $\Fq$.
\end{remark}

\begin{corollary}\label{cor:application}
\begin{enumerate}
\item 
For every algebraic torus $T$ over $K$, the trace of Frobenius 
  \[
\trFrob{T}:  \QCiso{\GN{T}} \to \Hom_{\text{ad}}(T(K),\EEx)
  \]
is an isomomorphism. 
\item The isomorphism $\trFrob{T}$ is functorial in the following sense: $T \mapsto \trFrob{T}$ is a natural transformation
from the additive functor
\[T \mapsto \QCiso{\GN{T}}\]
to the additive functor
\[
T \mapsto \Hom_{\text{ad}}(T(K),\EEx).
\]
\item
For every finite Galois extension $L/K$ and for every algebraic torus $T$ over $K$, the following diagram commutes:
\[
\begin{tikzcd}
\QCiso{\underline{\Res_{L/K}(T_L)}} \arrow[two heads]{r} \dar & \QCiso{\GN{T}} \dar\\
\Hom_{\text{ad}}(T(L), \EEx) \arrow[two heads]{r} & \Hom_{\text{ad}}(T(K), \EEx)
\end{tikzcd}
\]
\end{enumerate}
\end{corollary}

@@

\Clifton{Work on motivation.}

The techniques in this paper apply to all non-Archimdean local fields $K$ with residue field $k$, and to all algebraic tori $T$ over $K$, without placing any restrictions on the characteristic of $K$ or on the ramification of $T$. Because of this, it is natural to ask if we can compare quasicharacter sheaves on $T$

We recall the notion of $N$-congruent tori from \cite[\S 2]{CY}.  Suppose $T$ and $T'$
are tori over non-Archimedian local fields $K$ and $K'$, splitting over $L$ and $L'$ respectively.

\begin{definition}
 We say that $T$ and $T'$ are \emph{$N$-congruent} if there are isomorphisms
 \begin{align*}
  \alpha : \OL/\pi_K^N \OL &\to \OO{L'}/\pi_{K'}^N \OO{L'} \\
  \beta : \Gal(L/K) &\to \Gal(L'/K') \\
  \phi : X^*(T) &\to X^*(T')
 \end{align*}
 satisfying the following conditions:
 \begin{enumerate}
  \item $\alpha$ induces an isomorphism $\OK/\pi_K^N \OK \to \OO{K'}/\pi_{K'}^N \OO{K'}$,
  \item $\alpha$ is $\Gal(L/K)$-equivariant relative to $\beta$: $\alpha(\gamma x) = \beta(\gamma) \alpha(x)$
  for $\gamma \in \Gal(L/K)$ and $x \in \OL/\pi_K^N \OL$,
  \item $\phi$ is $\Gal(L/K)$-equivariant relative to $\beta$.
 \end{enumerate}
\end{definition}

One of the main results of \cite{CY} gives an isomorphism of group schemes between
$\TT_d$ and $\TT'_d$ assuming that $T$ and $T'$ are sufficiently congruent.
They define a quantity $h$ as the smallest integer so that $\pi^h$ lies in the
Jacobian ideal associated to a natural embedding of $T$ into an induced torus \cite[\S 8.1]{CY}.
They then prove the following theorem:

\begin{theorem}[{\cite[Thm. 8.5]{CY}}]
 Suppose that $T$ and $T'$ are $N$-congruent for $N > 3h$.  Then there is a canonical isomorphism
 \[
  \TT_{N-3h} \to \TT'_{N-3h}
 \]
 determined by $\alpha, \beta$ and $\phi$.
\end{theorem}

As an immediate consequence of this theorem and Theorem \ref{thm:application} we have the following corollary:

@@

@@

>>>>

\begin{theorem}\label{thm:application}
Let $d$ be a positive integer.
\begin{enumerate}
\item 
For every algebraic torus $T$ over $K$, the trace of Frobenius 
  \[
\trFrob{T,d}:  \QCiso{\GN{T}_d} \to \Hom_{< d}(T(K),\EEx)
  \]
is an isomomorphism. 
\item The isomorphism $\trFrob{T,d}$ is functorial in the following sense: $T \mapsto \trFrob{T,d}$ is a natural transformation
from the additive functor
\[T \mapsto \QCiso{\GN{T}_d}\]
to the additive functor
\[
T \mapsto \Hom_{<d}(T(K),\EEx).
\]
\item
For every finite Galois extension $L/K$ and for every algebraic torus $T$ over $K$, the following diagram commutes, where  . . . 
\[
\begin{tikzcd}
\QCiso{\Gr(\underline{\Res_{L/K}(T_L)}_d)} \arrow[two heads]{r} \dar & \QCiso{\Gr({T}_d)} \dar\\
\Hom_{<d}(T(L), \EEx) \arrow[two heads]{r} & \Hom_{<d}(T(K), \EEx)
\end{tikzcd}
\]
\end{enumerate}
\end{theorem}

@@

\begin{remark}
The Greenberg transform of the N\'eron model of $T$ is a commutative group scheme over $\Fq$. 
We will not use that fact in this paper. 
\end{remark}

\begin{corollary}\label{cor:application}
\begin{enumerate}
\item 
For every algebraic torus $T$ over $K$, the trace of Frobenius 
  \[
\trFrob{T}:  \QCiso{\GN{T}} \to \Hom_{\text{qc}}(T(K),\EEx)
  \]
is an isomomorphism. 
\item The isomorphism $\trFrob{T}$ is functorial in the following sense: $T \mapsto \trFrob{T}$ is a natural transformation
from the additive functor
\[T \mapsto \QCiso{\GN{T}}\]
to the additive functor
\[
T \mapsto \Hom_{\text{qc}}(T(K),\EEx).
\]
\item
For every finite Galois extension $L/K$ and for every algebraic torus $T$ over $K$, the following diagram commutes:
\[
\begin{tikzcd}
\QCiso{\Gr(\underline{\Res_{L/K}(T_L)})} \arrow[two heads]{r} \dar & \QCiso{\Gr({T})} \dar\\
\Hom_{\text{qc}}(T(L), \EEx) \arrow[two heads]{r} & \Hom_{\text{qc}}(T(K), \EEx)
\end{tikzcd}
\]
\end{enumerate}
\end{corollary}

@@

The techniques in this paper apply to all non-Archimdean local fields $K$ with residue field $k$, and to all algebraic tori $T$ over $K$, without placing any restrictions on the characteristic of $K$ or on the ramification of $T$. Because of this, it is natural to ask if we can compare quasicharacter sheaves on $T$ and on $T'$ under conditions when there is a natural comparison between characters of $T(K)$ and of $T'(K')$. This comparison of has been made precise in \cite{CY}, building on ideas of \cite{Deligne:limites}.

We recall the notion of $N$-congruent tori from  \cite{CY}*{\S 2}.  Suppose $T$ and $T'$
are tori over non-Archimedian local fields $K$ and $K'$, splitting over $L$ and $L'$ respectively.
Then $T$ and $T'$ are said to be \emph{$N$-congruent} if there are isomorphisms
 \begin{align*}
  \alpha : \OL/\pi_K^N \OL &\to \OO{L'}/\pi_{K'}^N \OO{L'} \\
  \beta : \Gal(L/K) &\to \Gal(L'/K') \\
  \phi : X^*(T) &\to X^*(T')
 \end{align*}
 satisfying the following conditions:
 \begin{enumerate}
  \item $\alpha$ induces an isomorphism $\OK/\pi_K^N \OK \to \OO{K'}/\pi_{K'}^N \OO{K'}$,
  \item $\alpha$ is $\Gal(L/K)$-equivariant relative to $\beta$: $\alpha(\gamma x) = \beta(\gamma) \alpha(x)$
  for $\gamma \in \Gal(L/K)$ and $x \in \OL/\pi_K^N \OL$,
  \item $\phi$ is $\Gal(L/K)$-equivariant relative to $\beta$.
 \end{enumerate}
\[
  \Hom_{<N}(T(K), \EEx) \to \Hom_{<N}(T'(K'),\EEx)
 \]

@@

\begin{proof}
One of the main results of \cite{CY} gives an isomorphism of group schemes between
$\TT_d$ and $\TT'_d$ assuming that $T$ and $T'$ are sufficiently congruent.
They define a quantity $h$ as the smallest integer so that $\pi^h$ lies in the
Jacobian ideal associated to a natural embedding of $T$ into an induced torus \cite{CY}*{\S 8.1}
They then prove the following theorem \cite{CY}*{Thm. 8.5}
 Suppose that $T$ and $T'$ are $N$-congruent for $N > 3h$.  Then there is a canonical isomorphism
 \[
  \TT_{N-3h} \to \TT'_{N-3h}
 \]
 determined by $\alpha, \beta$ and $\phi$.

As an immediate consequence of this theorem and Theorem \ref{thm:application} we have the following corollary:
\end{proof}

@@

\bib{Deligne:limites}{article}{
   author={Deligne, P.},
   title={Les corps locaux de caract\'eristique $p$, limites de corps locaux
   de caract\'eristique $0$},
%   language={French},
   conference={
      title={Representations of reductive groups over a local field},
   },
   book={
      series={Travaux en Cours},
      publisher={Hermann},
      place={Paris},
   },
   date={1984},
   pages={119--157},
%   review={\MR{771673 (86g:11068)}},
}
